\chapter{Abstract}
\label{chap:abstract}
\section{English}
Since the late 2000s more and more webpages when surfing the internet have added an option to use dark mode spending lots of time
developing this feature. This has led to a long standing debate on whether dark mode or bright mode are better. This 
thesis aims to clarify this debate around the readability and usability of texts in both modes.

The investigation is developed using an eye-tracker and a questionnaire to analyze the preference of users and the 
impact of these modes in the reading of the user. Thirty four subjects have been analyzed in a wiki-based and an e-commerce scenario.
After a through statistical analysis of the variables collected, the results didn't reveal significative results to dictaminate
that one mode has a benefit or disadvantage over the other in usability and legibility terms. However, the preference of users 
has been claryfied for users in bright mode over dark mode.
In conclusion, the results prove that this display options should be centered around preference and not objective advantages.
\newpage
\section{Español}
Desde finales de los años 2000, cada vez más páginas web han añadido la opción del modo oscuro, gatando tiempo y 
recursos en su desarrollo. Esto ha llevado a un largo debate sobre cual de los modos claro u oscuro es mejor. 
Este proyecto tiene como objetivo aclarar este debate en torno a la legibilidad y usabilidad de los textos en ambos modos. 

La investigación se desarrolla utilizando un eye-tracker y un cuestionario para analizar la preferencia de los usuarios y el impacto de estos
modos en la lectura del usuario. Treinta y cuatro sujetos han sido analizados en un escenario basado en wiki y otro de comercio electrónico. 
Después de un análisis exhaustivo de las variables colectadas, los resultados no revelaron resultados significativos para concluir que un 
modo tiene ventajas o desventajas en usabilidad y legibilidad sobre el otro. Sin embargo, se ha aclarado la preferencia de los usuarios por el 
modo claro sobre el modo oscuro. En conclusión, los resultados prueban que los modos deberían ser escogidos en función de la preferencia y no de ventajas objetivas.