\chapter{Data Analysis}
\label{chap:data_analysis}

\section{Data Preprocessing}
Data preprocessing is important for every investigation as it prevents the poisoning of the results 
with undesired outliers. In this case, there are two main sourrces of data, the tobii data and the 
google forms questionnaire. The data preprocessing included heavy manage of .tsv files and .csv files.
\subsection{Tobii Data}
The tobii data was preprocessed using the tool Tobii Pro Lab seen in chapter \ref{chap:methodology}.
The process consisted of importing the .tsv files and creating a pandas dataframe that would later be divided
into two different dataframes. The first dataframe would be the positive display polarity and the second 
dataframe would be the negative display polarity. This is possible thanks to the common column to all 
exported data called "Timeline". This specific column showed wich experiment was being performed at the 
moment of the recording. 

This separation was later divided for the different scenarios: Wikipedia webpage and Amazon webpage.
This is achieved thanks to the TOI (Time of Interest) column, which showed the scenes of the experiment.
\begin{lstlisting}[style=pythonstyle] 
def filter_by_polarity(df):
    '''
    Function to filter the DataFrame by polarity (claro and oscuro)

    There must be a column named Timeline with the values 'Timeline1-Mode1' for claro and 'Timeline2-Mode2' for oscuro
    '''
    claro = df[df['Timeline'] == 'Timeline1-Mode1'] 
    oscuro = df[df['Timeline'] == 'Timeline2-Mode2'] 
    return claro, oscuro
def separate_by_toi(df, toi):
    '''
    Function to obtain Patata, Mochila, Text and Entire Recording

    There must be a column named 'TOI' 
    '''
    df_toi = df[df['TOI'] == toi]
    return df_toi
tf = pd.read_csv('raw/TotalFixations.tsv', sep='\t')

tf_claro, tf_oscuro = filter_by_polarity(tf)


tf_claro_patata = separate_by_toi(tf_claro, 'PatataClaro')
tf_oscuro_patata = separate_by_toi(tf_oscuro, 'PatataOscuro')

tf_claro_mochila = separate_by_toi(tf_claro, 'MochilaClaro')
tf_oscuro_mochila = separate_by_toi(tf_oscuro, 'MochilaOscuro')

tf_claro_entire = separate_by_toi(tf_claro, 'Entire Recording')
tf_oscuro_entire = separate_by_toi(tf_oscuro, 'Entire Recording')
\end{lstlisting}

The data outliers were not taken into account for the analysis, as they were not 
considered to be a problem for the results and there were really few of them. Therefore no 
more preprocessing was needed for the tobii data.
\subsection{Google Forms Data}
The google forms data needed much preprocessing due to the fact of human intervention and a bug 
related to Tobii Pro Lab. Firstly the data was imported as a pandas dataframe from the CSV file
downloaded from the google spreadsheet generated for the form. 

The second step was to declutter the dataframe by removing all the duplicated rows. These duplicated rows
are generated by a bug in tobii pro lab that makes the explorer to send many requests rather than one, which 
ended up in generating an undefined number of duplicates per person. This was solved with the following code. 
Note that due to Latex nature the original column contains an accent in the "u" of número but the code blocks
do not allow it so it was changed to avoid the compiling error.

\begin{lstlisting}[style=pythonstyle]
    df = pd.read_csv('raw/Encuesta TFG.csv')
    df = df.drop_duplicates(subset=['Numero de participante'], keep='first').copy() 
\end{lstlisting}

Next, the age column needed detailed work as some people set their age and some others their birth date. This 
was solved with a function that checks if the value is a number, in such case it returns the number converted
to an integer. In the other case, it tries to parse the date with the pandas function `to\_datetime()` this can
result in a warning but it is not an issue. The warning explains that when using the days first format, a date is 
introduced starting with something else than the day.

\begin{lstlisting} [style=pythonstyle]
# Convertimos el string a fecha
df['Marca temporal'] = pd.to_datetime(df['Marca temporal'], dayfirst=True, errors='coerce')

def calculate_age(age_value, ref_ts):
    # Hay gente que puso la edad numerica solo
    age_num = pd.to_numeric(str(age_value).strip(), errors='coerce')
    if pd.notna(age_num):
        return int(age_num)

   # si puso fecha, la calculamos y comparamos con la fecha de referencia del csv (ref_ts)
   # Esto genera un warning pero da igual porq no es nada que afecte el codigo
    birth_ts = pd.to_datetime(str(age_value).strip(), dayfirst=True, errors='coerce') 
    if pd.notna(birth_ts) and pd.notna(ref_ts):
        years = ref_ts.year - birth_ts.year - ((ref_ts.month, ref_ts.day) < (birth_ts.month, birth_ts.day))
        return int(years)

# Llamamos a la funcion
df['Edad'] = [calculate_age(v, t) for v, t in zip(df['Edad'], df['Marca temporal'])]
    
\end{lstlisting}

In the case of the date being parsed, then the formula to calculate the age is the following:
\begin{equation}
    Age = \text{Current Year} - \text{Birth Year} - (1 \text{ or } 0)
\end{equation}
The final step of 1 or 0 checks whether the person has already had their birthday this year or not, 
in the case of having had the birthday, then the age is not reduced, otherwise it is reduced one year. 
After computing all the ages one subject did not provide the correct age information but rather gave the actual year.

Finally the last step was to preprocess the question regarding the type of vision of the subjects. This was done 
by hand as there were some unique cases that wouldn't make sense to be solved programming. This question will 
be further examined in the chapter \ref{chap:results}.
\section{Chosen Metrics}
The chosen metrics to analyze are based on the article "Eye tracking in human-computer interaction and usability 
research: Current status and future prospects"\cite{Poole2006}. These metrics are based on the eye-tracking data
and were analyzed for the different escenarios as long as with the whole recording. 
\subsection{Fixations}
Fixations are the moments when the eye stays still focused on a point of the screen. This moments can be as small
as a few milliseconds. There are a few metrics which are related to fixations as
shown below with their respective definition by Poole, Alex and Ball, Linden J.\cite{Poole2006}

\begin{table} [htb]
\caption{Fixation Metrics and their corresponding variables in the .tsv file}
    \begin{tabularx}{\textwidth}{|c|X|}
        \hline
        \textbf{Fixation Metric} & \textbf{Variable in .tsv file} \\
        \hline
        Gaze & Total\_duration\_of\_whole\_fixations \\
        \hline
        Total Number of Fixations & Number\_of\_whole\_fixations \\
        \hline
        Duration of First Fixation & Duration\_of\_first\_whole\_fixation \\
        \hline
        Average Duration of Fixations & Average\_duration\_of\_whole\_fixations \\
        \hline
    \end{tabularx}
\end{table}


\begin{itemize}
    \item \textbf{Gaze (sum of the duration of all fixations):} "Gaze is usually the sum of all fixation durations 
    within a prescribed area. It is best used to compare attention distributed between targets. It can 
    also be used as a measure of anticipation in situation awareness if longer gazes fall on an area of 
    interest before a possible event occurring."
    \begin{itemize}
        \item Although this metric is centered in Areas of Interest, in this thesis the 2 areas compared are
        positive display polarity and negative display polarity.
    \end{itemize}
    \item \textbf{Total Number of Fixations:} "More overall fixations indicate less efficient search 
    (perhaps due to sub-optimal layout of the interface)."
    \item \textbf{Duration of Fixation:} "A longer fixation duration indicates difficulty in extracting 
    information, or it means that the object is more engaging in some way."
    \begin{itemize}
        \item This will be analyzed both with the first fixation and with the average duration of the fixations.
    \end{itemize}
\end{itemize}
\subsection{Saccades}
Saccades are the time the eye is moving from one fixation to another, during this period the brain is not 
processing any information. They can provide information that supports the fixations or that can be useful 
in this thesis. These metrics will be explained briefly with the definitions given by Poole, Alex and Ball, 
Linden J.\cite{Poole2006}
\begin{itemize}
    \item \textbf{Saccade Amplitue:} "Larger saccades indicate more meaningful cues, as attention is drawn
     from a distance."
        \begin{itemize}
            \item The Saccade Amplitude is analyze through the average amplitude, 
            as well as the average peak velocity of the saccades.
        \end{itemize}
    \item \textbf{Number of Saccades:} "More saccades indicate more searching."
\end{itemize}

\begin{table} [htb]
\caption{Saccade Metrics and their corresponding variables in the .tsv file}
    \begin{tabularx}{\textwidth}{|c|X|}
        \hline
        \textbf{Saccade Metric} & \textbf{Variable in .tsv file} \\
        \hline
        Saccade Amplitude & Average\_peak\_velocity\_of\_saccades \\
        \hline
        Average Saccade Amplitude & Average\_amplitude\_of\_saccades \\
        \hline
        Amplitude of First Saccade & Amplitude\_of\_first\_saccade \\
        \hline
        Number of Saccades & Number\_of\_saccades \\
        \hline
    \end{tabularx}
\end{table}
TODO: Explain more saccades and there's missing one metric explanation in saccade
\section{Statistical Analysis}
\label{sec:statistical_analysis}
In this section the process of analyzing the data with statistical methods will be further explained.
This process consists of 3 main steps that are shared along all the studied data. Please, keep in mind that all three
statistical methods have the $\alpha$ value of 0,05 but relevant results under 0,1 will also be recorded. In adition, 
no extense explanation of the method will be given as this is out of the scope of the thesis.
\begin{enumerate}
    \item \textbf{Extract Basic Data:} Firstly the standard deviation, mean, maximum and minimum of each group
    is extracted to obtain the general shape. It will be later used in the representation of the results of the hypothesis.
    \item \textbf{Shapiro-Wilk Test\cite{shapiro-wilk}:} Secondly the data is analyzed with the Shapiro-Wilk test to check
    the normality of it. The null hypothesis of this test proves that the data follows a normal distribution, from now
    on, this statistic will be called $W$. In the following chapters $W_{PP}$ and  $W_{NP}$ will represent the Shapiro-Wilk test
    for Positive Polarity and Negative Polarity respectively.
    $$W = \frac{(\sum_{i=1}^n a_i x_{(i)})^2}{\sum_{i=1}^n (x_i - \bar{x})^2}$$
    \item \textbf{T-Student test\cite{t-test} or Mann-Whitney U\cite{mann-whitney} test\:} Finally, a comparison of both groups
    $W$ is compared. In the case of both datasets following a normal distribution, then the T-Student test is used, otherwise 
    the Mann-Whitney U test is used. Note that in both cases the $\alpha$ value is 0,05 but relevant results under 0,1 will also
    be recorded. 
    \subitem \textbf{T-Student test:} The T-Student test compares the means of two normall distributed groups. Being the
    null hypothesis that both groups have the same mean. Therefore it helps to check whether the difference between
    the groups is significant or not. In this study, the chosen variation is the independent samples test.
    $$ t = \frac{\bar{x}_1 - \bar{x}_2}{\sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}} $$
    \subitem \textbf{Mann-Whitney U test:} The Mann-Whitney U test is a non-parametric test which compares the medians of 
    two samples being the null hypothesis that both groups have the same median. Altough it is different from the T-Student
    results should be simmilar and it also provides a way to test data that is not normally distributed.
    $$ U_1 = n_1 n_2 + \frac{n_1 (n_1 + 1)}{2} - R_1 $$
    $$ U_2 = n_1 n_2 + \frac{n_2 (n_2 + 1)}{2} - R_2 $$
\end{enumerate}
